Efficient and Adjustable Run-Time Monitoring Using Dataflow-Guided Filtering

### 2014-09-11 dropping.tex
imported an authcomments package from Andrew Myers
Moved null filtering into Section 3.

% Run-time monitoring overview
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/monitoring_architecture.pdf}
    \vspace{-0.2in}
    \caption{Overview of run-time monitoring architecture.}
    \label{fig:monitoring.overview} 
    \vspace{-0.1in}
  \end{center}
\end{figure}

ED: monitoring core vs. monitor?

Figure~\ref{fig:monitoring.overview} shows an overview of the run-time monitoring
model that is assumed in this paper.  The \emph{main program} is a computation
task that performs the original function of the system and is run on the
\emph{main core}.  On certain events during the main program, such as the
execution of certain types of instructions, the \emph{monitoring core} performs a
series of \emph{monitoring operations}. The monitoring core operates in parallel to the
main core. These events are referred to as \emph{monitoring events}. Depending
on the type of monitoring event, different monitoring operations are
executed. Information about monitoring events are sent to the monitoring core and buffered in a FIFO structure to decouple the
running of the main core and the monitoring core. If the FIFO is full, then the main
core is forced to stall on a monitoring event until a FIFO entry becomes
available. These stalls are a major source of
overhead because the monitoring core may take several cycles to process a single event
from the main core. 
%We refer to these stalls and other overheads, such as
%contention for shared resources, as \emph{monitoring overheads}. 
If the monitoring core
detects an inconsistent or undesired behavior in the monitoring events, then
an error is detected.

ED: describe reg/mem meta-data both in the figure and in the text.

@@

% Example monitor
\begin{figure*}
  \begin{center}
    \includegraphics[]{figs/example_full.pdf}
    \vspace{-0.2in}
    \caption{Example of array bounds check monitor.}
    \label{fig:monitoring.example_full}
    \vspace{-0.1in}
  \end{center}
\end{figure*}

Figure~\ref{fig:monitoring.example_full} shows an example pseudo-code segment,
its assembly level instructions, and the corresponding monitoring operations
for an array bounds check monitor. 
First, an array {\tt x} is allocated using {\tt malloc}. 
As {\tt malloc} returns the array's address in a register, the monitor
associates base and bounds metadata with the corresponding metadata. 
Next, pointer {\tt y} is set to point to the middle of array {\tt x}. 
At the assembly level, a register {\tt r2} is set to array {\tt x}'s address plus an offset.
The monitor propagates the metadata of the original pointer in {\tt r1} to the
metadata of {\tt r2}. This is to ensure that pointer {\tt y} is not used to exceed the array's bounds.
Finally, both pointers {\tt x} and {\tt y} are used to perform a write. 
This requires setting register {\tt r3} to the write value. When this happens,
{\tt r3}'s metadata is also reset to NULL in case it previously stored a
pointer. In both cases, the monitor checks whether the address that is written
to is within the bounds stored in the metadata. In the first case, {\tt x+12}
is within the original array's bounds. No error is raised and the metadata of
{\tt r3} is propagated to the metadata of the store address. If {\tt r3} were a
pointer, than this would allow a future instruction to load the pointer and use
it to access the corresponding array or data structure. In the second case,
{\tt y+12} corresponds to {\tt x+20} which is not within the array's bounds and
the monitor will raise an error.

@@

The key cause of false positives is dropping monitoring operations that update
metadata. Dropping monitoring operations that check for an error (such as check
found for line 4 in Figure~\ref{fig:monitoring.example_full} can only cause
false negatives but will never cause false positives. On the other hand,
skipping monitoring operations that update metadata can lead to false positives
and false negatives.
Essentially, when an update operation is skipped, we can no longer trust the
corresponding metadata. Thus, our general approach for preventing false positives is to mark these metadata as
invalid. Furthermore, this invalidation information is propagated in the same
way that metadata is. Figure~\ref{fig:dropping.example_invalid}
shows an example of associating invalid flags with metadata. In the example, no
monitoring operations are dropped. However, suppose that the monitoring for
line 1 is dropped. In this case, {\tt rf\_invalid[1]} will be marked as {\tt
true}. Thus, when line 3 is reached, the monitoring event is dropped
since {\tt rf\_invalid[1]} is marked as true. Note that in this case, line 2
also propagates this invalid flag to {\tt rf\_invalid[2]} and causes the check
performed on line 4 to also be dropped. This is necessary because an error
would have been raised even if the access on line 4 was in bounds.

@@

>>>>

% Run-time monitoring overview
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/monitoring_architecture.pdf}
    \vspace{-0.2in}
    \caption{Overview of run-time monitoring architecture.}
    \label{fig:dropping.overview} 
    \vspace{-0.1in}
  \end{center}
\end{figure}

\ED{Should we use monitoring core or just monitor?}

Figure~\ref{fig:dropping.overview} shows an overview of the run-time monitoring
model that is assumed in this paper.  The \emph{main program} is a computation
task that performs the original function of the system and is run on the
\emph{main core}.  On certain events during the main program, such as the
execution of certain types of instructions, the \emph{monitoring core} performs a
series of \emph{monitoring operations}. The monitoring core operates in parallel to the
main core. These events are referred to as \emph{monitoring events}. Depending
on the type of monitoring event, different monitoring operations are
executed. Information about monitoring events are sent to the monitoring core and buffered in a FIFO structure to decouple the
running of the main core and the monitoring core. If the FIFO is full, then the main
core is forced to stall on a monitoring event until a FIFO entry becomes
available. These stalls are a major source of
overhead because the monitoring core may take several cycles to process a single event
from the main core. 
%We refer to these stalls and other overheads, such as
%contention for shared resources, as \emph{monitoring overheads}. 
If the monitoring core
detects an inconsistent or undesired behavior in the monitoring events, then
an error is detected.

\ED{describe reg/mem meta-data both in the figure and in the text.}

@@

% Example monitor
\begin{figure*}
  \begin{center}
    \includegraphics[]{figs/example_full.pdf}
    \vspace{-0.2in}
    \caption{Example of array bounds check monitor.}
    \label{fig:dropping.example_full}
    \vspace{-0.1in}
  \end{center}
\end{figure*}

Figure~\ref{fig:dropping.example_full} shows an example pseudo-code segment,
its assembly level instructions, and the corresponding monitoring operations
for an array bounds check monitor. 
First, an array {\tt x} is allocated using {\tt malloc}. 
As {\tt malloc} returns the array's address in a register, the monitor
associates base and bounds metadata with the corresponding metadata. 
Next, pointer {\tt y} is set to point to the middle of array {\tt x}. 
At the assembly level, a register {\tt r2} is set to array {\tt x}'s address plus an offset.
The monitor propagates the metadata of the original pointer in {\tt r1} to the
metadata of {\tt r2}. This is to ensure that pointer {\tt y} is not used to exceed the array's bounds.
Finally, both pointers {\tt x} and {\tt y} are used to perform a write. 
This requires setting register {\tt r3} to the write value. When this happens,
{\tt r3}'s metadata is also reset to NULL in case it previously stored a
pointer. In both cases, the monitor checks whether the address that is written
to is within the bounds stored in the metadata. In the first case, {\tt x+12}
is within the original array's bounds. No error is raised and the metadata of
{\tt r3} is propagated to the metadata of the store address. If {\tt r3} were a
pointer, than this would allow a future instruction to load the pointer and use
it to access the corresponding array or data structure. In the second case,
{\tt y+12} corresponds to {\tt x+20} which is not within the array's bounds and
the monitor will raise an error.

@@

The key cause of false positives is dropping monitoring operations that update
metadata. Dropping monitoring operations that check for an error (such as check
found for line 4 in Figure~\ref{fig:dropping.example_full} can only cause
false negatives but will never cause false positives. On the other hand,
skipping monitoring operations that update metadata can lead to false positives
and false negatives.
Essentially, when an update operation is skipped, we can no longer trust the
corresponding metadata. Thus, our general approach for preventing false positives is to mark these metadata as
invalid. Furthermore, this invalidation information is propagated in the same
way that metadata is. Figure~\ref{fig:dropping.example_invalid}
shows an example of associating invalid flags with metadata. In the example, no
monitoring operations are dropped. However, suppose that the monitoring for
line 1 is dropped. In this case, {\tt rf\_invalid[1]} will be marked as {\tt
true}. Thus, when line 3 is reached, the monitoring event is dropped
since {\tt rf\_invalid[1]} is marked as true. Note that in this case, line 2
also propagates this invalid flag to {\tt rf\_invalid[2]} and causes the check
performed on line 4 to also be dropped. This is necessary because an error
would have been raised even if the access on line 4 was in bounds.

@@

\subsection{Filtering Null Metadata}
\label{sec:dropping.null_filtering}

% Null filtering code example
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/example_null_filtering.pdf}
    \vspace{-0.2in}
    \caption{Example of using information about null metadata to filter monitoring events.}
    \label{fig:dropping.example_null} 
    \vspace{-0.1in}
  \end{center}
\end{figure}

One way to reduce the number of monitoring events that must be handled by the
monitor is to filter out monitoring events that correspond to operating on null
or uninitialized metadata. Null metadata typically corresponds to events that
are not relevant to the monitor. For example, Hardbound
\cite{hardbound-asplos08} filters out operations on non-pointer (i.e., no base
and bounds metadata) instructions since it is not relevant to array bounds
checking. More recently, FADE \cite{fade-hpca14} has been proposed a general
hardware module to perform this null metadata filtering for a variety of
monitoring schemes.

Figure~\ref{fig:dropping.example_null} shows an example of how this null filtering operates.
Here, the main core's code has been slightly modified from
Figure~\ref{fig:dropping.example_full} and {\tt r2} on line 2 is no longer set as an array pointer.
Without null metadata filtering, the instructions for lines 2 and 4 would be
forwarded to the monitor since the system does not know whether {\tt r2}
contains a pointer address or not. However, if we use a flag to mark {\tt r2}
as null when it is loaded with a constant, then we can propagate this null
information and filter out monitoring for lines 2 and 4.

The operations performed by null filtering are almost identical to the
operations needed for invalidation shown in
Figure~\ref{fig:dropping.example_invalid}.
Thus, we can easily enable this null metadata filtering on our architecture by
simply extending the dataflow flags to be two bits wide. One bit is used to
keep track of invalidation information while the second bit is used to keep
track of null information. All flags are initialized to null and the filter
decision table is extended with the propagation and filtering decision rules
for null metadata filtering.
